### 2017-01-29 latex/dataMaid_toyData.Rmd
Updated article

---
dataMaid: yes
title: toyData
subtitle: "Autogenerated data summary from dataMaid"
date: "`r Sys.Date()`"
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
# Data cleaning summary
The dataset examined has the following dimensions:

------------------------------
Feature                 Result
--------------------- --------
Number of rows              15

Number of variables          6
------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

----------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$

Identify levels with < 6 obs.                          $\times$    $\times$   $\times$

Identify case issues                                   $\times$    $\times$   $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$

Identify outliers                                                                        $\times$   $\times$             $\times$ 
----------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Variable list
## **var1**

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                    factor

Number of missing obs.      2 (13.33 %)

Number of unique values               2

Mode                              "red"
---------------------------------------

\emini
\bminitwo
```{r Var-1-var1, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
qplot(x = structure(c(2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
1L, 1L, 1L), .Label = c("blue", "red"), class = "factor", na.action = structure(14:15, class = "omit")), 
    geom = "bar", xlab = "", main = "var1")
```
\emini

- Note that the following levels have at most five observations: \"blue\".

\fullline

## **var2**

\bminione

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                numeric

Number of missing obs.      3 (20 %)

Number of unique values            8

Median                           4.5

1st and 3rd quartiles        1.75; 6

Min. and max.                 1; 999
------------------------------------

\emini
\bminitwo
```{r Var-2-var2, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
qplot(x = c(1, 1, 1, 2, 2, 6, 6, 6, 999, 4, 82, 5), geom = "histogram", 
    xlab = "", main = "var2", bins = 20)
```
\emini

- The following suspected missing value codes enter as regular values: \"999\", \"NaN\".

- Note that the following possible outlier values were detected: \"82\", \"999\".

\fullline

## **var3**

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                factor

Number of missing obs.      0 (0 %)

Number of unique values           7

Mode                            "a"
-----------------------------------

\emini
\bminitwo
```{r Var-3-var3, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
qplot(x = structure(c(3L, 3L, 3L, 3L, 3L, 4L, 4L, 4L, 5L, 5L, 
5L, 2L, 1L, 6L, 7L), .Label = c(" ", ".", "a", "b", "c", "other", 
"OTHER"), class = "factor"), geom = "bar", xlab = "", main = "var3")
```
\emini

- The following suspected missing value codes enter as regular values: \" \", \".\".

- The following values appear with prefixed or suffixed white space: \" \".

- Note that the following levels have at most five observations: \" \", \".\", \"a\", \"b\", \"c\", \"other\", \"OTHER\".

- Note that there might be case problems with the following levels: \"other\", \"OTHER\".

\fullline

## **var4**

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                   numeric

Number of missing obs.          0 (0 %)

Number of unique values              15

Median                             0.33

1st and 3rd quartiles       -0.62; 0.66

Min. and max.                -2.21; 1.6
---------------------------------------

\emini
\bminitwo
```{r Var-4-var4, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
qplot(x = c(-0.626453810742332, 0.183643324222082, -0.835628612410047, 
1.59528080213779, 0.329507771815361, -0.820468384118015, 0.487429052428485, 
0.738324705129217, 0.575781351653492, -0.305388387156356, 1.51178116845085, 
0.389843236411431, -0.621240580541804, -2.2146998871775, 1.12493091814311
), geom = "histogram", xlab = "", main = "var4", bins = 20)
```
\emini

- Note that the following possible outlier values were detected: \"1.12\", \"1.51\", \"1.6\".

\fullline

## **var5**

* The variable is a key (distinct values for each observation).

\fullline

## **var6**

* The variable only takes one (non-missing) value: \"Irrelevant\". The variable contains 0 \% missing observations.

\fullline

This report was created by dataMaid v0.9.2.