### 2017-12-21 docs/articles/dataMaid_artData.Rmd
Added package site

---
dataMaid: yes
title: artData
subtitle: "Autogenerated data summary from dataMaid"
date: 2017-12-21 01:29:36
output: html_document
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
```{r visualFunctions, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggAggHist <- getFromNamespace("ggAggHist", "dataMaid")
ggAggBarplot <- getFromNamespace("ggAggBarplot", "dataMaid")
```
# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations        200

Number of variables            11
---------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character   factor    labelled   numeric   integer   logical    Date   
----------------------------------------------------- ----------- --------- ---------- --------- --------- --------- ---------
Identify miscoded missing values                        &times;    &times;   &times;    &times;   &times;

Identify prefixed and suffixed whitespace               &times;    &times;   &times;

Identify levels with < 6 obs.                           &times;    &times;   &times;

Identify case issues                                    &times;    &times;   &times;

Identify misclassified numeric or integer variables     &times;    &times;   &times;

Identify non-trailing colons                            &times;    &times;   &times;

Identify outliers                                                                       &times;   &times;             &times; 
------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

---------------------------------------------------------------------------------------------
&nbsp;              Variable class     # unique values   Missing observations  Any problems? 
------------------- ---------------- ----------------- ---------------------- ---------------
[ArtistID]          character                      179                 0.00 %     &times;

[ArtistName]        character                      179                 0.00 %     &times;

[NoOfMiddlenames]   numeric                          4                 0.00 %     &times;

[Title]             character                      200                 0.00 %     &times;

[Year]              integer                        149                 0.00 %     &times;

[Location]          character                       98                 1.50 %     &times;

[Continent]         factor                           3                 0.00 %     &times;

[Width]             numeric                        164                 5.00 %     &times;

[Height]            numeric                        165                 5.00 %     &times;

[Media]             character                       28                 5.00 %     &times;

[Movement]          character                       86                 9.00 %     &times;    
---------------------------------------------------------------------------------------------

# Variable list
## ArtistID

* Warning: This variable seems to contain ID codes.

---

## ArtistName

<div class = "row">
<div class = "col-lg-8">

---------------------------------------------
Feature                                Result
------------------------- -------------------
Variable type                       character

Number of missing obs.                0 (0 %)

Number of unique values                   179

Mode                        "Diego Velazquez"

No. zeros                                   0
---------------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-2-ArtistName, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:179, .Label = c("Adolph von Menzel", 
"Alberto Giacometti", "Albrecht Altdorfer", "Albrecht Durer", 
"Alexej von Jawlensky", "Andrea Mantegna", "Andrew Wyeth", "Annibale Carracci", 
"Anthony van Dyck", "Antoine Watteau", "Antonio Canaletto", "Arnold Bocklin", 
"Arshile Gorky", "August Macke", "Auguste Renoir", "Bartolome Esteban Murillo", 
"Benozzo Gozzoli", "Caravaggio", "Carl Spitzweg", "Caspar David Friedrich", 
"Claude Monet", "Correggio", "David Wilkie", "Diego Velazquez", 
"Dong Qichang", "Duccio di Buoninsegna", "Edgar Degas", "Edouard Manet", 
"Edvard Munch", "Edward Hopper", "Egon Schiele", "El Greco", 
"Elisabeth Vigee-Lebrun", "Emil Nolde", "Emilio Vedova", "Erich Heckel", 
"Ernst Ludwig Kirchner", "Eugene Delacroix", "Ferdinand Hodler", 
"Fernand Khnopff", "Fernand Leger", "Francis Bacon", "Francis Picabia", 
"Francisco Goya", "Francisco de Zurbaran", "Francois Boucher", 
"Franz Kline", "Franz Marc", "Franz von Lenbach", "Gabriele Munter", 
"Geertgen tot Sint Jans", "Georg Baselitz", "George Grosz", "George Stubbs", 
"Georges Braque", "Georges Seurat", "Georges de la Tour", "Georgia O'Keeffe", 
"Giacomo Balla", "Giorgione", "Giotto", "Giovanni Battista Tiepolo", 
"Giovanni Bellini", "Giulio Romano", "Giuseppe Arcimboldo ", 
"Grant Wood", "Gustave Caillebotte", "Gustave Courbet", "Hannah Hoch", 
"Hans Baldung Grien", "Hans Holbein the Younger", "Hans Memling", 
"Hans von Marees", "Helen Frankenthaler", "Hendrick Avercamp", 
"Henri Matisse", "Henri Rousseau", "Henry Fuseli", "Henry Raeburn", 
"Hieronymus Bosch", "Honore Daumier", "Hugo van der Goes", "Hyacinthe Rigaud", 
"Ilya Repin", "Jackson Pollock", "Jacob Izaaksoon van Ruisdael", 
"Jacopo Tintoretto", "Jacques-Louis David", "James Ensor", "James McNeill Whistler", 
"Jan Matsys", "Jan Vermeer", "Jan van Eyck", "Jasper Johns, Jr.", 
"Jean Auguste Dominique Ingres", "Jean Metzinger", "Jean Simeon Chardin", 
"Joachim Patinir", "Joan Miro", "Johannes Itten", "John Constable", 
"John Everett Millais", "Josef Albers", "Joseph Wright of Derby", 
"Juan Gris", "Kangra-School", "Karl Friedrich Schinkel", "Kazimir Malevich", 
"Konrad Klapheck", "Kurt Schwitters", "Laszlo Moholy-Nagy", "Leonardo da Vinci", 
"Lorenzo Lotto", "Louis or Antoine Le Nain", "Lovis Corinth", 
"Luca Signorelli", "Lucas Cranach the Elder", "Lucio Fontana", 
"Lyonel Feininger", "Marc Chagall", "Marcel Duchamp", "Mark Rothko", 
"Mary Cassatt", "Masaccio", "Matthias Grunewald", "Max Beckmann", 
"Max Ernst", "Mikhail Nesterov", "Morris Louis", "Nicolas Poussin", 
"Niklaus Manuel", "Oskar Kokoschka", "Oskar Schlemmer", "Otto Dix", 
"Pablo Picasso", "Paolo Uccello", "Paolo Veronese", "Paul Cezanne", 
"Paul Gauguin", "Paul Klee", "Paul Nash", "Peter Paul Rubens", 
"Philipp Otto Runge", "Piero della Francesca", "Piet Mondrian", 
"Pieter Bruegel the Elder", "Pieter Saenredam", "Raphael", "Reginald Marsh", 
"Rembrandt", "Rene Magritte", "Richard Lindner", "Robert Delaunay", 
"Rogier van der Weyden", "Roy Lichtenstein", "Salvador Dali", 
"Sandro Botticelli", "Sebastiano Ricci", "Sonia Delaunay-Terk", 
"Stanley Spencer", "Stefan Lochner", "Tawaraya S?tatsu", "Theodore Gericault", 
"Thomas Eakins", "Thomas Gainsborough", "Titian", "Umberto Boccioni", 
"Victor Pasmore", "Vincent van Gogh", "Vittore Carpaccio", "Wassily Kandinsky", 
"Wilhelm Leibl", "Wilhelm von Kaulbach", "Wilhelm von Kobell", 
"William Holman Hunt", "William McTaggart", "William Turner", 
"Winslow Homer", "Wolf Vostell"), class = "factor"), y = c(1L, 
1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
2L, 1L, 1L, 1L, 1L, 1L, 3L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 2L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
2L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 2L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 
3L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 2L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
1L, 1L)), .Names = c("x", "y"), row.names = c(NA, -179L), class = "data.frame"), 
    vnam = "ArtistName")
```
</div>
</div>

- The following values appear with prefixed or suffixed white space: \"Giuseppe Arcimboldo \".

- Note that the following levels have at most five observations: \"Adolph von Menzel\", \"Alberto Giacometti\", \"Albrecht Altdorfer\", \"Albrecht Durer\", \"Alexej von Jawlensky\", \"Andrea Mantegna\", \"Andrew Wyeth\", \"Annibale Carracci\", \"Anthony van Dyck\", \"Antoine Watteau\" (169 additional values omitted).

---

## NoOfMiddlenames

* Note that this variable is treated as a factor variable below, as it only takes a few unique values.

<div class = "row">
<div class = "col-lg-8">

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               numeric

Number of missing obs.      0 (0 %)

Number of unique values           4

Mode                            "0"

No. zeros                       157
-----------------------------------

</div>
<div class = "col-lg-4">
```{r Var-3-NoOfMiddlenames, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
mosaicplot(structure(c(157L, 29L, 13L, 1L), .Dim = 4L, .Dimnames = structure(list(
    v = c("0", "1", "2", "3")), .Names = "v"), class = "table"), 
    main = "NoOfMiddlenames", xlab = "")
```
</div>
</div>

- Note that the following levels have at most five observations: \"3\".

---

## Title

* The variable is a key (distinct values for each observation).

---

## Year

<div class = "row">
<div class = "col-lg-8">

-----------------------------------------
Feature                            Result
------------------------- ---------------
Variable type                     integer

Number of missing obs.            0 (0 %)

Number of unique values               149

Median                             1851.5

1st and 3rd quartiles       1627.75; 1914

Min. and max.                  1150; 1968

No. zeros                               0

Mean                              1765.73
-----------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-5-Year, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
prettierHistHelper(v = c(1886L, 1914L, 1531L, 1470L, 1930L, 1914L, 
1768L, 1912L, 1894L, 1914L, 1939L, 1950L, 1524L, 1576L, 1931L, 
1515L, 1917L, 1720L, 1961L, 1926L, 1907L, 1915L, 1958L, 1888L, 
1911L, 1948L, 1150L, 1957L, 1515L, 1919L, 1517L, 1559L, 1938L, 
1928L, 1914L, 1566L, 1913L, 1955L, 1936L, 1559L, 1965L, 1937L, 
1965L, 1911L, 1951L, 1910L, 1637L, 1930L, 1490L, 1843L, 1887L, 
1904L, 1550L, 1656L, 1530L, 1922L, 1881L, 1506L, 1448L, 1894L, 
1751L, 1871L, 1815L, 1934L, 1638L, 1920L, 1496L, 1968L, 1917L, 
1952L, 1749L, 1942L, 1877L, 1863L, 1944L, 1852L, 1877L, 1645L, 
1717L, 1492L, 1505L, 1955L, 1701L, 1635L, 1520L, 1780L, 1913L, 
1958L, 1891L, 1460L, 1732L, 1595L, 1911L, 1460L, 1823L, 1630L, 
1661L, 1786L, 1914L, 1889L, 1899L, 1500L, 1630L, 1508L, 1950L, 
1880L, 1479L, 1912L, 1633L, 1598L, 1942L, 1635L, 1465L, 1670L, 
1903L, 1528L, 1450L, 1900L, 1529L, 1456L, 1487L, 1914L, 1586L, 
1936L, 1835L, 1896L, 1637L, 1891L, 1459L, 1846L, 1640L, 1954L, 
1916L, 1911L, 1852L, 1621L, 1893L, 1500L, 1885L, 1965L, 1762L, 
1914L, 1310L, 1851L, 1806L, 1858L, 1565L, 1666L, 1660L, 1789L, 
1596L, 1524L, 1434L, 1824L, 1953L, 1306L, 1800L, 1824L, 1839L, 
1460L, 1819L, 1665L, 1927L, 1893L, 1808L, 1784L, 1799L, 1533L, 
1897L, 1890L, 1634L, 1898L, 1425L, 1562L, 1862L, 1510L, 1965L, 
1563L, 1736L, 1944L, 1882L, 1912L, 1793L, 1936L, 1922L, 1920L, 
1915L, 1660L, 1610L, 1911L, 1910L, 1745L, 1630L, 1852L, 1804L, 
1630L, 1912L, 1886L, 1867L, 1903L), vnam = "Year")
```
</div>
</div>

- Note that the following possible outlier values were detected: \"1958\", \"1961\", \"1965\", \"1968\".

---

## Location

<div class = "row">
<div class = "col-lg-8">

----------------------------------------------
Feature                                 Result
------------------------- --------------------
Variable type                        character

Number of missing obs.               3 (1.5 %)

Number of unique values                     97

Mode                        "National Gallery"

No. zeros                                    0
----------------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-6-Location, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:97, .Label = c("Alfred Stieglitz Collection, Carl Van Vechten Gallery", 
"Alte Nationalgalerie", "Alte Pinakothek", "Art Institute of Chicago", 
"Brancacci Chapel", "Centre Pompidou", "Christie's Auction House", 
"Cleveland Museum of Art", "Courtauld Institute of Art, University of London", 
"Detroit Institute of Arts", "Fogg Museum", "Forberg Collection, Albertina", 
"Freer Gallery of Art", "Galleria dell'Accademia Carrara", "Gallerie dell'Accademia", 
"Gemaeldegalerie Alte Meister", "Gemaldegalerie", "Gemaldegalerie Alte Meister", 
"Gemeentemuseum Den Haag", "Guggenheim Museum", "Hamburger Kunsthalle", 
"Harvard Art Museum", "Hermitage Museum", "Hill-Stead Museum", 
"Kroller-Muller Museum", "Kunsthalle (Museum of Art)", "Kunsthaus Zurich", 
"Kunsthistorisches Museum", "Kunstmuseum Basel", "Kunstsammlung Nordrhein-Westfalen", 
"Kunthistorisches Museum", "Lenbach House State Gallery", "Leopold Hoesch Museum", 
"Louvre Museum", "Manchester Art Gallery", "Mauritshuis", "Medici Riccardi Palace", 
"Memlingmuseum, Sint-Janshospitaal", "Menard Art Museum", "Metropolitan Museum of Art", 
"Mr. and Mrs. Leigh B. Block", "Musee National d'Art Moderne", 
"Musee Royaux des Beaux-Arts", "Musee d'Orsay", "Musee d'Unterlinden", 
"Musee des Beaux-Arts de Nantes", "Museo Civico", "Museo Reina Sofia", 
"Museo del Novecento", "Museo del Prado", "Museum Folkwant", 
"Museum Gemaldegalerie der Staatlichen Museen", "Museum Ludwig", 
"Museum of Contemporary Art", "Museum of Fine Arts", "Museum of Modern Art", 
"National Gallery", "National Gallery of Art", "National Gallery of Scotland", 
"National Gallery, Galleria degli Uffizi, Louvre Museum", "Nationalgalerie, Staatliche Museen", 
"Nationalgalerie, Staatliche Museen zu Berlin", "Neue Pinakothe", 
"New Pinakothek", "Norton Simon Museum", "Osterreichische Galerie Belvedere", 
"Palazzo Ducale", "Peggy Guggenheim Collection", "Pennsylvania Academy of the Fine Arts", 
"Philadelphia Museum of Art", "Phillips Collections", "Pompidou Centre", 
"Private Collection", "Rijksmuseum", "Royal Collection, Windsor Castle", 
"Royal Museums of Fine Arts of Beligum", "Samuel H. Kress Collection, National Gallery of Art", 
"San Francisco Museum of Modern Art", "Santo Tome", "Scrovegni Chapel", 
"Staatliche Museen", "Stadtische Galerie im Lenbachhaus", "State Russian Museum", 
"Stedelijk Museum", "Szepmuveszeti Muzeum", "Tate Britain", "The Frick Collection", 
"Toledo Museum of Art", "Trustees of the Grosvenor Estate", "Uffizi Gallery", 
"Victoria & Albert Museum", "Villa Coredellina", "Wallraf-Richartz Museum", 
"Walters Art Museum", "Whitney Museum of American Art", "Wildenstein Collection", 
"von Zezschwitz (for sale)"), class = "factor"), y = c(1L, 1L, 
2L, 8L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 4L, 2L, 
1L, 1L, 4L, 1L, 1L, 1L, 1L, 2L, 3L, 4L, 6L, 3L, 4L, 2L, 1L, 12L, 
1L, 1L, 1L, 1L, 1L, 4L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 7L, 
1L, 1L, 2L, 1L, 1L, 8L, 13L, 6L, 3L, 1L, 1L, 1L, 4L, 1L, 1L, 
1L, 1L, 6L, 1L, 1L, 1L, 1L, 2L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
1L, 2L, 1L, 1L, 4L, 1L, 1L, 1L, 1L, 1L, 1L, 3L, 1L, 1L, 1L, 1L
)), .Names = c("x", "y"), row.names = c(NA, -97L), class = "data.frame"), 
    vnam = "Location")
```
</div>
</div>

- Note that the following levels have at most five observations: \"Alfred Stieglitz Collection, Carl Van Vechten Gallery\", \"Alte Nationalgalerie\", \"Alte Pinakothek\", \"Brancacci Chapel\", \"Centre Pompidou\", \"Christie's Auction House\", \"Cleveland Museum of Art\", \"Courtauld Institute of Art, University of London\", \"Detroit Institute of Arts\", \"Fogg Museum\" (79 additional values omitted).

---

## Continent

<div class = "row">
<div class = "col-lg-8">

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                 factor

Number of missing obs.       0 (0 %)

Number of unique values            3

Mode                        "Europe"

No. zeros                          0
------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-7-Continent, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
mosaicplot(structure(c(2L, 143L, 55L), .Dim = 3L, .Dimnames = structure(list(
    v = c("Asia", "Europe", "North America")), .Names = "v"), class = "table"), 
    main = "Continent", xlab = "")
```
</div>
</div>

- Note that the following levels have at most five observations: \"Asia\".

---

## Width

<div class = "row">
<div class = "col-lg-8">

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type                 numeric

Number of missing obs.       10 (5 %)

Number of unique values           163

Median                         122.45

1st and 3rd quartiles       77; 198.3

Min. and max.               10.7; 990

No. zeros                           0

Mean                           168.46
-------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-8-Width, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
prettierHistHelper(v = c(308, 76.52, 35.5, 242, 62.4, 57, 244, 
107.5, 141, 97.1, 96.3, 207, 193, 167, NA, 77, 390, 199, 439.4, 
71.1, 88.9, 73, 279.4, 65.5, 128.6, 121.3, 150.8, 68, 589, 144, 
14.7, 204.5, 101.6, 73, 250, 51, 37, 153.87, 248.92, 112.9, 121.9, 
776, 61, 151.4, 170.4, 99.5, NA, 76.2, 28, 93, 92, 129.2, 112, 
276, 191, 100, 172.7, 88, 40, 87, 73, 117.5, NA, 13.9, 87.5, 
79, 389, 120, 160, 297.8, 119.04, 152.4, 46.6, 190, 119.3, 111.8, 
276, 159.5, 194, 32, 26, 60, 194, 209, 161, 17.2, 10.7, 297.8, 
358, 200, 126, 148.1, 73, 144, 111.8, 257, 77, 84, 73, 45, 80, 
48.7, 96, 175, 100.3, 130, 331.5, 100, 107, 196.2, 292.1, 214, 
127, 110, NA, 64, 116, 250.8, 120.3, 957, 278.5, 220, 360, 27, 
123, 150, 229.2, 66.1, 93, 705, 81, 131.2, NA, 95, 205, 123.5, 
174, 389, 190, 300, NA, 50, 45, 109.5, 141, 325, 162, 166.5, 
180, 422, 126.5, 85.4, 62, 354, 182.9, 185, 190, 96.7, 44.6, 
NA, 716, 179.39, 548.6, 73.5, 305, 25, 15.2, 81.5, 200.7, 183, 
367, 178, 597, 162, 108, 112, NA, 990, 61, 222, 77, 111.5, 135, 
101.6, 48.1, 60, 97.8, 117.5, 108.6, NA, 50.17, 35, 369.9, 130, 
61, 46.4, 129.9, 70, 205, NA), vnam = "Width")
```
</div>
</div>

- Note that the following possible outlier values were detected: \"10.7\", \"13.9\", \"14.7\", \"15.2\", \"17.2\", \"25\", \"26\", \"27\", \"28\", \"32\" (3 additional values omitted).

---

## Height

<div class = "row">
<div class = "col-lg-8">

-----------------------------------------
Feature                            Result
------------------------- ---------------
Variable type                     numeric

Number of missing obs.           10 (5 %)

Number of unique values               164

Median                                113

1st and 3rd quartiles       73.12; 168.75

Min. and max.                   12.3; 666

No. zeros                               0

Mean                               134.51
-----------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-9-Height, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
prettierHistHelper(v = c(207.6, 54.61, 50.4, 147, 74.3, 88, 183, 
78.5, 120.5, 130.4, 129.6, 105, 175, 146, NA, 59.5, 260, 118.5, 
262.3, 43.8, 116.2, 92, 195.6, 81, 160.7, 81.9, 37.5, 48.3, 336, 
90, 30.3, 187, 67.9, 91.8, 250, 67, 48.5, 107.32, 198.12, 113.2, 
121.9, 349, 61, 192.1, 130.5, 107, NA, 101.6, 42, 62, 67, 169.5, 
73.5, 318, 152, 131, 129.9, 113, 51, 66, 59, 82.6, NA, 17.1, 
63, 104.5, 365, 200, 150, 220, 69.08, 84.1, 60.3, 130.5, 94.2, 
76.2, 212.2, 113.3, 129, 50, 35, 73, 277, 173, 151.5, 12.3, 18.1, 
270.8, 203, 225, 77, 88.3, 100, 77, 87.6, 148, 91, 105, 57, 51, 
120, 67.1, 71, 108.5, 81.3, 66, 176, 86, 60, 141, 199.39, 154, 
81, 130, NA, 76.8, 168, 210.5, 158.4, 182, 172.5, 181, 460, 35, 
92.7, 50, 131.2, 100.3, 67, 585, 93, 195.4, NA, 70, 142, 101.9, 
96.5, 220, 145, 250, NA, 60, 43, 76.4, 131, 207, 117, 121.5, 
140, 323, 100, 125.7, 66, 419, 152.4, 200, 97, 126.9, 36.2, NA, 
491, 136.21, 274.3, 91, 202, 30, 21.8, 115, 129.5, 122, 307, 
195, 247, 117, 108, 168, NA, 666, 66, 94, 113, 111, 169, 76.2, 
63.8, 81, 125.7, 98.5, 121.3, NA, 60.96, 64, 166, 163, 74, 29.4, 
193, 70, 255, NA), vnam = "Height")
```
</div>
</div>

- Note that the following possible outlier values were detected: \"460\", \"491\", \"585\", \"666\".

---

## Media

<div class = "row">
<div class = "col-lg-8">

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                 character

Number of missing obs.         10 (5 %)

Number of unique values              27

Mode                        "oil paint"

No. zeros                             0
---------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-10-Media, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:27, .Label = c("acrylic paint", 
"aquatint, etching", "collage", "collage, oil paint", "egg tempera", 
"enamel", "encaustic, oil paint, and collage on fabric", "fresco", 
"gouache and graphite", "gouache-painted paper", "ink and color", 
"ink, color, gold and silver", "oil and metallic paint", "oil paint", 
"oil paint and charcoal", "oil paint and collage", "oil paint and magna", 
"oil paint and pastel", "oil paint, tempera, pastel", "oil piant", 
"pastel", "screen photography, transparent paint, silkscreen", 
"tempera", "varnished tempera", "watercolor", "watercolor, pen and ink", 
"wood cut"), class = "factor"), y = c(2L, 1L, 1L, 1L, 2L, 1L, 
1L, 4L, 1L, 1L, 1L, 1L, 1L, 147L, 2L, 1L, 1L, 1L, 1L, 2L, 1L, 
1L, 9L, 1L, 2L, 1L, 2L)), .Names = c("x", "y"), row.names = c(NA, 
-27L), class = "data.frame"), vnam = "Media")
```
</div>
</div>

- Note that the following levels have at most five observations: \"acrylic paint\", \"aquatint, etching\", \"collage\", \"collage, oil paint\", \"egg tempera\", \"enamel\", \"encaustic, oil paint, and collage on fabric\", \"fresco\", \"gouache and graphite\", \"gouache-painted paper\" (15 additional values omitted).

---

## Movement

<div class = "row">
<div class = "col-lg-8">

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.       18 (9 %)

Number of unique values            85

Mode                        "Baroque"

No. zeros                           0
-------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-11-Movement, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:85, .Label = c("Abstract Artist", 
"Abstract Expressionist", "Abstract Expressionist:Color Field", 
"Abstract Expressionist:Neo-Dada:Pop Art", "American Modernist", 
"American Realist", "Analytic Cubist", "Baroque", "Bauhaus", 
"British Expressionist", "Classicist:Baroque", "Constructivist", 
"Cubism", "Cubist", "Dada", "Dada:New Objectivity", "Dada:Surrealist", 
"De Stijl", "Dutch Golden Age", "Dutch Golden Age:Baroque", "Dutch and Flemish Renaissance", 
"Early Netherlandish", "Early Netherlandish Renaissance", "Early Renaissance", 
"English Romantic", "Expressionist", "Expressionist:Abstract Artist", 
"Expressionist:Bauhaus:Surrealist", "Expressionist:Surrealist", 
"Fauvist", "Flemish", "French Baroque", "French Impressionist", 
"Futurism", "Geometric abstraction", "German Expressionist", 
"German Realist", "German Renaissance", "German Romantic", "High Renaissance", 
"High Renaissance:Mannerist", "Impressionist", "Impressionist:Expressionist", 
"Italian Modernist", "Italian Renaissance", "Late Gothic", "Mannerist", 
"Mechanistic Cubist", "Merz", "Neo-Expressionist", "Neo-Impresionist:Divisionist:Fauvist:Cubist", 
"Neoclassicist", "Neue Sachlichkeit", "Northern Renaissance", 
"Orphist", "Orphist:Cubist:Expressionist", "Pop Artist", "Post-Impressionist", 
"Post-Impressionist:Neo-Impressionist:Pointillist", "Pre-Raphaelite", 
"Realism", "Realist", "Realist:Impressionist", "Realist:Surrealist:Pop Artist", 
"Regionalist", "Renaissance", "Rimpa school", "Rococo", "Rococo:Neoclassicist", 
"Romantic", "Russian Expressionist", "Russian Symbolist", "Scottish landscape painter", 
"Sienese school:Gothic style", "Social Realism", "Spatialist", 
"Suprematist", "Surrealism", "Surrealism:Dada", "Surrealist", 
"Surrealist:Expressionist", "Surrealist:Expressionist:Cubist:Formalist", 
"Symbolist", "Tonalist", "Video Artist:Installation:Happening:Fluxus"
), class = "factor"), y = c(1L, 5L, 1L, 1L, 1L, 1L, 1L, 14L, 
1L, 1L, 2L, 1L, 4L, 2L, 1L, 1L, 3L, 1L, 1L, 4L, 3L, 1L, 2L, 2L, 
1L, 11L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 3L, 3L, 3L, 
6L, 1L, 4L, 1L, 1L, 8L, 2L, 3L, 1L, 1L, 1L, 1L, 2L, 1L, 7L, 1L, 
1L, 1L, 6L, 1L, 2L, 1L, 5L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 9L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 3L, 1L, 1L)), .Names = c("x", 
"y"), row.names = c(NA, -85L), class = "data.frame"), vnam = "Movement")
```
</div>
</div>

- Note that the following levels have at most five observations: \"Abstract Artist\", \"Abstract Expressionist\", \"Abstract Expressionist:Color Field\", \"Abstract Expressionist:Neo-Dada:Pop Art\", \"American Modernist\", \"American Realist\", \"Analytic Cubist\", \"Bauhaus\", \"British Expressionist\", \"Classicist:Baroque\" (68 additional values omitted).

- Note that the following values include colons: \"Abstract Expressionist:Color Field\", \"Abstract Expressionist:Neo-Dada:Pop Art\", \"Classicist:Baroque\", \"Dada:New Objectivity\", \"Dada:Surrealist\", \"Dutch Golden Age:Baroque\", \"Expressionist:Abstract Artist\", \"Expressionist:Bauhaus:Surrealist\", \"Expressionist:Surrealist\", \"High Renaissance:Mannerist\" (12 additional values omitted).

---

Report generation information:

 *  Created by Claus Thorn Ekstrøm.

 *  Report creation time: Thu Dec 21 2017 01:29:36

 *  dataMaid v1.0.1 [Pkg: 2017-11-13 from local (ekstroem/dataMaid@NA)]

 *  R version 3.4.3 (2017-11-30).

 *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.1).

 *  Function call: `makeDataReport(data = artData, output = "html", preChecks = c("isKey", 
    "isSingular", "isSupported", "isID"), replace = TRUE, openResult = FALSE, 
    summaries = setSummaries(character = defaultCharacterSummaries(add = "countZeros"), 
        factor = defaultFactorSummaries(add = "countZeros"), 
        labelled = defaultLabelledSummaries(add = "countZeros"), 
        numeric = defaultNumericSummaries(add = c("countZeros", 
            "meanSummary")), integer = defaultIntegerSummaries(add = c("countZeros", 
            "meanSummary")), logical = defaultLogicalSummaries(add = c("meanSummary"))), 
    visuals = setVisuals(factor = "mosaicVisual", numeric = "prettierHist", 
        integer = "prettierHist", Date = "prettierHist"), checks = setChecks(character = defaultCharacterChecks(add = "identifyColons"), 
        factor = defaultFactorChecks(add = "identifyColons"), 
        labelled = defaultLabelledChecks(add = "identifyColons")))`