### 2018-02-08 docs/CSP2018/slides/module2/module2.Rmd

---
title: "Cleaning Up the Data Cleaning Process"
subtitle: "Module 2: Data Screening with dataMaid"
author: "Claus Thorn Ekstrøm and Anne Helby Petersen<br>UCPH Biostatistics"
date: "CSP, Portland, OR, Feb. 15th, 2018<br>.footnotesize[[Slides/homepage](https://ekstroem.github.io/dataMaid/CSP2018/index.html)]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ['default', 'ku.css', 'animate.min.css']
    seal: true
    nature:
      ratio: '16:9'
      slideNumberFormat: '%current%' 
      highlightStyle: ir-black
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(pics/manbeer.jpg)
background-size: 100%
class: middle, center

```{r setup, include=FALSE}
library(knitr)
options(htmltools.dir.version = FALSE, cache=TRUE)
opts_chunk$set(comment = NA, prompt=TRUE)
#opts_chunk$set(dev.args=list(bg="transparent"), fig.width=15, fig.height=7)
source("kutheme.R")
```

# The RESCueH project

---

class: center

# Timeline followback (TLFB)

```{r, echo=FALSE}
DF <- data.frame(day1=sample(1:25, 6),
                 day2=c(NA, NA, sample(1:25, 3), 88),
                 day3=c(NA, 99, 40, sample(1:25, 3)))
head(DF, 4)
```

---

class: center

# Timeline followback (TLFB)

```{r, echo=FALSE}
head(DF, 6)
```

---

background-image: url(pics/mau.png)
background-size: 80%
class: center

# Monthly Alcohol units

```{r echo=FALSE,message=FALSE, warning=FALSE, error=FALSE}
library(extrafont)
loadfonts()
library(ggplot2)
library(ggthemr)
claus <- define_palette(swatch = c('#eeeeee', '#FFFACD','#901A1E', 'lightblue', 'green', 'blue', 'brown', 'purple', 'yellow'),
  gradient = c(lower = 'red', upper = 'green'),
                        background = "#222222",
                        text = c("#eeeeee", "#eeeeee"),
                        line = c("#ee1e1e", "#6e6e6e"), # Axis lines
                        gridline = "#434343") # Gridlines

ggthemr(claus, layout="scientific", type = 'outer', spacing=1.2)

tlfb <- data.frame(units=dget("tlfb.txt"))

p <- ggplot(tlfb, aes(units)) + geom_dotplot() +
#     ggtitle("Baseline alcohol") +
     xlab("Units") + ylab("Frequency") +
     theme(text=element_text(size=24, family="Comic Sans MS"),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank())

theme_xkcd <- theme(

#  panel.background = element_rect(fill="transparent"),
#  plot.background = element_rect(fill="transparent"),

  #axis.ticks = element_line(colour=NA),
##  panel.grid = element_line(colour="white"),
  #axis.text.y = element_text(colour=NA),
##  axis.text.x = element_text(colour="black"),
  text = element_text(size=36, family="Humor Sans")
)
## p + theme_xkcd

# Rotate the dot plot
# p + coord_flip()

#p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point(size=3) +
#  ggtitle("Fuel Efficiency of 32 Cars") +
#  xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
#  theme(text=element_text(size=16, family="Comic Sans MS"))
#p
```

---

# Reproducible research

What **didn't** we check?

--

- Need **experts in relevant field**
- Merge existing databases into megadatabases
- New technologies revive old data

---

class: center

```{r message=FALSE, echo=FALSE}
library(dataMaid)
data(toyData)
```

.small[
```{r message=FALSE, echo=FALSE}
toyData
```
]

---

class: middle

# dataMaid

```{r eval=FALSE}
library(dataMaid)
data(toyData)
makeDataReport(toyData) #<<
```

Documentation to be **read** and **evaluated** by a human.

See [github.com/ekstroem/dataMaid](github.com/ekstroem/dataMaid) for more info. Stable version on CRAN.

---

background-image: url(pics/flowchart2.png)
background-size: 100%
class: center

# `dataMaid` flowchart

```{r eval=TRUE,echo=FALSE, warning=FALSE, out.width='100%', out.height='60%'}
library(DiagrammeR)
library(DiagrammeRsvg)
library(svglite)
library(rsvg)

grViz("
  digraph {

graph[layout = neato,
      overlap=false,
      rankdir = 'LR', 
      bgcolor='#000000',
      sep=.2]

node[fontsize=20,
      fillcolor = DimGray,
      fontcolor = White,
      fillcolor='#222222',
      style=filled,
      color=White]

edge [color = White, penwidth=4]

  'data frame'  -> 'summarize'
    subgraph detail {
       rank='same';
       'summarize' -> 'visualize' -> 'check';
       'check' -> 'summarize'; }
'check' -> '.Rmd / render';
  }")
```

---

# Part 1: Data cleaning summary

```{r res1, fig.align='center', echo=FALSE, out.width='100%'}
knitr::include_graphics('pics/summ.png', dpi = NA)
```

---

background-image: url(pics/miss.png)
background-size: 100%

# Part 2: Summary table

---

background-image: url(pics/out1.png)
background-size: 100%

# Part 3: Variable list

---

background-image: url(pics/out2.png)
background-size: 100%

---

# `dataMaid` common arguments

.small[

| Argument      | Description   |
| ------------- |:--------------------------------| 
| `mode`        | Tasks to perform. `c("summarize", "visualize", "check")` is default |
| `replace`      | Logical. Should existing dataMaid reports be overwritten? Default `FALSE`   | 
| `output`      | Output format. Choices are `"pdf"`, `"html"`, "`"word"` | 
| `onlyProblematic`      | Logical. Show only variable with problems. Default `FALSE`  |
| `maxProbVals`  | Maximum number of unique values printed. Positive int or `Inf` (default 10)     |

]

---

class: inverse, middle

# Exercise 2

1.  Load the `dataMaid` package
2.  Use `makeDataReport()` to generate a data report
3.  Screen the `bigPresidentData` data for errors we did not find before.

Hunt for errors!

---

class: center

.small[
```{r message=FALSE, echo=FALSE}
toyData
```
]

---

# Using `dataMaid` interactively

.footnotesize[
```{r}
check(toyData$events)
check(toyData$events, 
        numericChecks = "identifyMissing")
```
]

???

check er de funktioner, der bliver checket for

Kan sætte specifikke (også i makeDataReport)

Vis, at det er en liste med 2 elementer, og den form, som de har.

---

# Overview of `check` functions

.footnotesize[
```{r eval=FALSE,results='markup'}
allCheckFunctions()
```
]

.footnotesize[

```
----------------------------------------------------------------------------------
name                 description                     classes                      
-------------------- -------------------------------------------------------------
identifyCaseIssues   Identify case issues            character, factor

identifyLoners       Identify levels with < 6 obs.   character, factor

identifyMissing      Identify miscoded missing       character, Date, factor,     
                     values                          integer, labelled, logical,  
                                                          numeric                      
```                                                       
]

and more

---

# Interactive `dataMaid` - visualizations

.pull-left[
.footnotesize[
```{r, eval=FALSE}
visualize(toyData$events)
```

Can also check the available `visual` functions

```{r eval=FALSE}
allVisualFunctions()
```
]
]

.pull-right[
```{r, eval=TRUE, echo=FALSE}
visualize(toyData$events)  # Visualize variable
```
]

---

# Interactive `dataMaid` - summaries

.footnotesize[
```{r, eval=TRUE}
summarize(toyData$events)
```
]

---

# Using dataMaid interactively

.small[
```
> allSummaryFunctions()

-------------------------------------------------------------
name         description          classes                    
------------ -------------------- ---------------------------
centralValue Compute median       character, Date, factor,   
             or mode              integer, labelled, logical,
                                  numeric

countMissing Compute ratio of     character, Date, factor,   
             missing obs.         integer, labelled, logical,
                                  numeric

minMax       Find min and max     integer, numeric, Date     
             values                                                   
quartiles    Compute 1st and 3rd quartiles Date, integer, numeric

uniqueValues Count number of unique values character, Date, factor,   
                                           integer, labelled, logical,
                                           numeric

variableType Data class of variable        character, Date, factor,   
                                           integer, labelled, logical,
                                           numeric                    
----------------------------------------------------------------------

```
]

---

# Extending `dataMaid`

Custom check, visual, or summary functions.

Few requirements --- input and output formats.

Easiet to work with a **template system** and modify one of those.

Check vignette `vignette("extending_dataMaid")` for detailed instructions. Or the exercises!

---

## Custom `summaryFunction` - template

.footnotesize[
```
mySummaryFunction <- function(v, ...) {
  val <- [ result of whatever summary we are doing ]
  res <- [ properly escaped version of val ]        
  summaryResult(list(feature = "[Feature name]", 
                     result = res,
                     value = val))
}
```
]

Example (centralValue for numeric/integer)

.footnotesize[
```{r eval=FALSE}
function (v, maxDecimals = 2) {
    v <- na.omit(v)
    val <- median(v)
    summaryResult(list(feature = "Median", result = round(val, 
        maxDecimals), value = val))
}
```
]

---

## Custom `checkFunction` - template

.small[
```{r, eval=TRUE}
isSSN <- function(v, nMax = NULL, ...) {
  out <- list(problem = FALSE, 
              message = "", 
              problemValues=NULL)
  if (class(v) %in% c("character", "factor", 
                      "labelled")) {
    if (any(grep("\\d{3}-\\d{2}-\\d{4}", v))) {
      out$problem <- TRUE
      out$message <- "Warning: Seems to contain SSNs."
      out$problemValues <- "Will not show"
    }  
  }
  out 
}
```
]
---

## Using the function

.footnotesize[
```{r}
DF <- data.frame(ids=c("111-22-3333","123-45-6789", 
                       "111-22-3333"),
                 id2=c("111223333", "123456789", "4728491283"), 
                 stringsAsFactors=FALSE)

check(DF, characterChecks = c("isSSN"))
```

]

---

class: inverse

# Exercise 2b/2c

How to tailor `dataMaid` to work with *your* dataset:

*   Work interactively with `dataMaid` (exercise 2b)
*   Create new custom summary, check and visualization functions (exercise 2c)

Pick whatever you want. Or jump back and forth.