### 2018-05-10 docs/eRum2018/slides/module3/pics/module4.Rmd
Added eRum slides+exercises

---
title: "Reproducible data screening and quality control"
subtitle: "Module 3: Time to clean and other options"
author: "Claus Thorn Ekstrøm and Anne Helby Petersen<br>UCPH Biostatistics"
date: "eRum, Budapest, May 14th, 2018<br>.footnotesize[[Slides/homepage](https://ekstroem.github.io/dataMaid/eRum2018/index.html)]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ['default', 'ku.css', 'animate.min.css']
    seal: true
    nature:
      ratio: '16:9'
      slideNumberFormat: '%current%' 
      highlightStyle: ir-black
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "js/zoom.js"
---

```{r setup, include=FALSE}
library(knitr)
options(htmltools.dir.version = FALSE, cache=TRUE)
opts_chunk$set(comment = NA, prompt=TRUE)
#opts_chunk$set(dev.args=list(bg="transparent"), fig.width=15, fig.height=7)
source("kutheme.R")
library(dataMaid)
toyData <- as.data.frame(toyData)
```

#Summarizing the errors

* What did we find so far?
* What didn't we find?

---

#Data cleaning

.center[
```{r cleanpic, echo=FALSE, out.width="60%"}
knitr::include_graphics("pics/datacleaning.jpg")
```
]

Not the best term ... and should not be unsupervised

---

#Data cleaning in R

In an R-script:

1. Make a copy of the dataset. 
2. Use indexing to locate the problem in the data.
3. Overwrite the faulty value with a correct one - if you know it - or `NA` to mark that information is missing in this spot. 
4. Save the copy of the "cleaned" data in a **new** file.

---

#Selection - rows/observations

Two systems for selecting observations in `data.frame`s in R:

By index (row number) or using a logical vector.

```{r}
(tD <- head(toyData, 3))  
```

---

#Selection - rows/observations

Four equivalent ways to get the second line of `tD`:
```{r, eval = FALSE}
tD[2, ] #indexing
tD[c(FALSE, TRUE, FALSE), ] #manual logical vector 
tD[tD$id == 2, ] #informative logical vector
tD %>% filter(id==2)  # Using tidyverse
```

```{r, echo = FALSE}
tD[tD$id == 2, ] #informative logical vector
```

---

#Selection - rows/observations

Use informative logical vectors as much as possible!

```{r}
tD

#Mark non-positive change as missing:
tD[tD$change > 0, "change"] <- NA
```

---

#Selection - columns/variables

*ALWAYS* use variable names.

```{r, eval = FALSE}
#readable, informative code:
tD[tD$change > 0, "change"] <- NA

# Indexing by numbers easily becomes 
# a source of error by itself:
tD[tD$change > 0, 4] <- NA
```

---

background-image: url(pics/structure.png)
background-size: 30%
background-position: right

## Finishing up after cleaning

Should now have<br>
a cleaned dataset<br>
that can form the<br>
basis for future<br>
analyses.<br>

With documentation<br>
of how we got<br>
there!

---

# Create codebook

Produce a summary document for subsequent analyses.

.footnotesize[
```{r eval=FALSE}
makeCodebook(presidentData)
```
]

Add label (similar to `labelled` package) or extra information

.footnotesize[
```{r}
pD <- presidentData
attr(pD$presidencyYears, "label") <- 
  "Full years as president"
```
]

.footnotesize[
```{r}
attr(pD$birthday, "shortDescription") <- 
  "Dates are stored in YYYY-MM-DD format"
```
]

---

class: inverse

# Exercise 3

Correct the errors you have found so far.

Make sure to make the cleaning process reproducible.

Remember **rules 1 and 2**!

Create the final codebook with additional information about some of the variables.

```{r eval=FALSE}
makeCodebook(myCleanedData)
```

---

##Row-wise or column-wise checks?

```{r colrow1, echo=FALSE, out.width="75%", fig.align = "center"}
knitr::include_graphics("pics/colrow1.png")
```

---

##Row-wise or column-wise checks?

```{r colrow2, echo=FALSE, out.width="75%", fig.align = "center"}
knitr::include_graphics("pics/colrow2.png")
```

---

##Row-wise or column-wise checks?

```{r colrow3, echo=FALSE, out.width="75%", fig.align = "center"}
knitr::include_graphics("pics/colrow3.png")
```

---

##Row-wise *and* column-wise constraints!

* `dataMaid` performs class dependent checks for each variable in a dataset, one at a time (column-wise)
  + Pros: Easy to document what was (not) done, let's you get started without a lot of prior knowledge, easy to share with collaborators
  + Shortcomings: Generally cannot detect internal consistency issues or use non-class dependent variable constraints

---

##Row-wise *and* column-wise constraints!

An R-packages that performs row-wise checks: `validate`. Check out the talk on Wednesday @ 14.50 by Edwin de Jonge:

**validatetools - resolve and simplify contradictive or redundant data validation rules**

Note: Different use of the term "validation" - no longer about format, type and range, but used as synonym to "check".

---

class: middle, center

# Thank you!

Please grab hold of us here or via email

.pull-left[Anne<br>[ahpe@sund.ku.dk](mailto:ahpe@sund.ku.dk)]  .pull-right[Claus<br>[ekstrom@sund.ku.dk](mailto:ekstrom@sund.ku.dk)]