### 2019-09-12 testdata_dateprob.rmd
Label ordering is now preserved for factor visualizations. Fixes #46.

---
dataMaid: yes
title: testData
subtitle: "Autogenerated data summary from dataMaid"
date: 2019-07-10 10:40:23
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library("ggplot2")
library("pander")
```

```{r 'visualFunctions', echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggAggHist <- getFromNamespace("ggAggHist", "dataMaid")
ggAggBarplot <- getFromNamespace("ggAggBarplot", "dataMaid")
```

# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations         15

Number of variables            15
---------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

---------------------------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   haven labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$      $\times$      $\times$   $\times$             $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$      $\times$

Identify levels with < 6 obs.                          $\times$    $\times$   $\times$      $\times$

Identify case issues                                   $\times$    $\times$   $\times$      $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$      $\times$

Identify outliers                                                                                         $\times$   $\times$             $\times$ 
---------------------------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

-------------------------------------------------------------------------------------------------
&nbsp;                  Variable class     # unique values   Missing observations  Any problems? 
----------------------- ---------------- ----------------- ---------------------- ---------------
[charVar]               character                        5                 6.67 %    $\times$

[factorVar]             factor                           5                 0.00 %    $\times$

[numVar]                numeric                         10                 0.00 %

[intVar]                integer                         10                 0.00 %

[boolVar]               logical                          3                20.00 %

[keyVar]                character                       15                 0.00 %    $\times$

[emptyVar]              numeric                          1                 0.00 %    $\times$

[numOutlierVar]         numeric                         15                 0.00 %    $\times$

[smartNumVar]           numeric                          2                 0.00 %

[cprVar]                character                        5                 0.00 %    $\times$

[cprKeyVar]             character                       15                 0.00 %    $\times$

[miscodedMissingVar]    character                       14                 0.00 %    $\times$

[misclassifiedNumVar]   factor                          12                 0.00 %    $\times$

[dateVar]               Date                             4                 0.00 %

[labelledVar]           labelled                         3                13.33 %    $\times$    
-------------------------------------------------------------------------------------------------

# Variable list
## charVar

\bminione

--------------------------------------
Feature                         Result
------------------------- ------------
Variable type                character

Number of missing obs.      1 (6.67 %)

Number of unique values              4

Mode                               "a"
--------------------------------------

\emini
\bminitwo
```{r 'Var-1-charVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:4, .Label = c("a", 
"b", "c", "d"), class = "factor"), y = c(6L, 4L, 2L, 2L)), class = "data.frame", row.names = c(NA, 
-4L)), vnam = "charVar")
```

\emini

- Note that the following levels have at most five observations: \"b\", \"c\", \"d\".

\fullline

## factorVar

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                factor

Number of missing obs.      0 (0 %)

Number of unique values           5

Mode                            "a"

Reference category              999
-----------------------------------

\emini
\bminitwo
```{r 'Var-2-factorVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:5, .Label = c("999", 
"a", "b", "c", "d"), class = "factor"), y = c(1L, 6L, 4L, 2L, 
2L)), class = "data.frame", row.names = c(NA, -5L)), vnam = "factorVar")
```

\emini

- The following suspected missing value codes enter as regular values: \"999\".

- Note that the following levels have at most five observations: \"999\", \"b\", \"c\", \"d\".

\fullline

## numVar

\bminione

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                numeric

Number of missing obs.       0 (0 %)

Number of unique values           10

Median                             5

1st and 3rd quartiles       1.5; 6.5

Min. and max.                  1; 10
------------------------------------

\emini
\bminitwo
```{r 'Var-3-numVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(xmin = c(0, 2, 4, 6, 8), xmax = c(2, 
4, 6, 8, 10), ymin = c(0, 0, 0, 0, 0), ymax = c(5L, 2L, 4L, 2L, 
2L)), class = "data.frame", row.names = c(NA, -5L)), vnam = "numVar")
```

\emini

\fullline

## intVar

\bminione

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                integer

Number of missing obs.       0 (0 %)

Number of unique values           10

Median                             5

1st and 3rd quartiles       1.5; 6.5

Min. and max.                  1; 10
------------------------------------

\emini
\bminitwo
```{r 'Var-4-intVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(xmin = c(0, 2, 4, 6, 8), xmax = c(2, 
4, 6, 8, 10), ymin = c(0, 0, 0, 0, 0), ymax = c(5L, 2L, 4L, 2L, 
2L)), class = "data.frame", row.names = c(NA, -5L)), vnam = "intVar")
```

\emini

\fullline

## boolVar

\bminione

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                logical

Number of missing obs.      3 (20 %)

Number of unique values            2

Mode                          "TRUE"
------------------------------------

\emini
\bminitwo
```{r 'Var-5-boolVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("FALSE", 
"TRUE"), class = "factor"), y = c(4L, 8L)), class = "data.frame", row.names = c(NA, 
-2L)), vnam = "boolVar")
```

\emini

\fullline

## keyVar

* The variable is a key (distinct values for each observation).

\fullline

## emptyVar

* The variable only takes one (non-missing) value: \"1\". The variable contains 0 \% missing observations.

\fullline

## numOutlierVar

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type                 numeric

Number of missing obs.        0 (0 %)

Number of unique values            15

Median                              8

1st and 3rd quartiles       4.5; 11.5

Min. and max.                  1; 100
-------------------------------------

\emini
\bminitwo
```{r 'Var-8-numOutlierVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(xmin = c(0, 20, 40, 60, 80), 
    xmax = c(20, 40, 60, 80, 100), ymin = c(0, 0, 0, 0, 0), ymax = c(14L, 
    0L, 0L, 0L, 1L)), class = "data.frame", row.names = c(NA, 
-5L)), vnam = "numOutlierVar")
```

\emini

- Note that the following possible outlier values were detected: \"100\".

\fullline

## smartNumVar

* Note that this variable is treated as a factor variable below, as it only takes a few unique values.

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               numeric

Number of missing obs.      0 (0 %)

Number of unique values           2

Mode                            "1"

Reference category                0
-----------------------------------

\emini
\bminitwo
```{r 'Var-9-smartNumVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("0", 
"1"), class = "factor"), y = 7:8), class = "data.frame", row.names = c(NA, 
-2L)), vnam = "smartNumVar")
```

\emini

\fullline

## cprVar

\bminione

-----------------------------------------
Feature                            Result
------------------------- ---------------
Variable type                   character

Number of missing obs.            0 (0 %)

Number of unique values                 5

Mode                        "010188-1350"
-----------------------------------------

\emini
\bminitwo
```{r 'Var-10-cprVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:5, .Label = c("010188-1350", 
"010290-0219", "010291-2373", "020187-3282", "040506-9660"), class = "factor"), 
    y = c(3L, 3L, 3L, 3L, 3L)), class = "data.frame", row.names = c(NA, 
-5L)), vnam = "cprVar")
```

\emini

- Note that the following levels have at most five observations: \"010188-1350\", \"010290-0219\", \"010291-2373\", \"020187-3282\", \"040506-9660\".

\fullline

## cprKeyVar

* The variable is a key (distinct values for each observation).

\fullline

## miscodedMissingVar

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            14

Mode                              "9"
-------------------------------------

\emini
\bminitwo
```{r 'Var-12-miscodedMissingVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:13, .Label = c("", 
"-", "-inf", "-Inf", ".", "9", "inf", "Inf", "na", "Na", "NA", 
"nan", "NAN"), class = "factor"), y = c(1L, 1L, 1L, 1L, 1L, 2L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L)), class = "data.frame", row.names = c(NA, 
-13L)), vnam = "miscodedMissingVar")
```

\emini

- The following suspected missing value codes enter as regular values: \"\", \"-\", \"-inf\", \"-Inf\", \".\", \"9\", \"inf\", \"Inf\", \"na\", \"Na\" (4 additional values omitted).

- Note that the following levels have at most five observations: \"\", \"-\", \"-inf\", \"-Inf\", \".\", \"9\", \"inf\", \"Inf\", \"na\", \"Na\" (4 additional values omitted).

- Note that there might be case problems with the following levels: \"-inf\", \"-Inf\", \"inf\", \"Inf\", \"na\", \"Na\", \"NA\", \"nan\", \"NaN\", \"NAN\".

\fullline

## misclassifiedNumVar

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                factor

Number of missing obs.      0 (0 %)

Number of unique values          12

Mode                            "1"

Reference category                1
-----------------------------------

\emini
\bminitwo
```{r 'Var-13-misclassifiedNumVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:12, .Label = c("1", 
"2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"), class = "factor"), 
    y = c(2L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 2L, 1L, 1L, 1L)), class = "data.frame", row.names = c(NA, 
-12L)), vnam = "misclassifiedNumVar")
```

\emini

- Note that the following levels have at most five observations: \"1\", \"10\", \"11\", \"12\", \"2\", \"3\", \"4\", \"5\", \"6\", \"7\" (2 additional values omitted).

- Note: The variable consists exclusively of numbers and takes a lot of different values. Is it perhaps a misclassified numeric variable?

\fullline

## dateVar

\bminione

--------------------------------------------------
Feature                                     Result
------------------------- ------------------------
Variable type                                 Date

Number of missing obs.                     0 (0 %)

Number of unique values                          4

Mode                                  "1997-05-10"

Min. and max.               1997-05-10; 2013-04-02

1st and 3rd quartiles       1997-05-10; 2005-12-10
--------------------------------------------------

\emini
\bminitwo
```{r 'Var-14-dateVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(xmin = structure(c(9000, 10000, 
11000, 12000, 13000, 14000, 15000), class = "Date"), xmax = structure(c(10000, 
11000, 12000, 13000, 14000, 15000, 16000), class = "Date"), ymin = c(0, 
0, 0, 0, 0, 0, 0), ymax = c(10L, 0L, 0L, 0L, 3L, 0L, 2L)), row.names = c(NA, 
-7L), class = "data.frame"), vnam = "dateVar")
```

\emini

\fullline

## labelledVar

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                  labelled

Number of missing obs.      2 (13.33 %)

Number of unique values               2

Mode                              "Joe"
---------------------------------------

\emini
\bminitwo
```{r 'Var-15-labelledVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("Joe", 
"Peter"), class = "factor"), y = c(8L, 5L)), class = "data.frame", row.names = c(NA, 
-2L)), vnam = "labelledVar")
```

\emini

- Note that the following levels have at most five observations: \"Peter\".

\fullline

Report generation information:

 *  Created by annennenne (username: `zms499`).

 *  Report creation time: Wed Jul 10 2019 10:40:24

 *  Report was run from directory: `C:/Users/zms499/Documents/P-kopier/dataMaid`

 *  dataMaid v1.3.0 [Pkg: 2019-07-10 from local]

 *  R version 3.5.3 (2019-03-11).

 *  Platform: x86_64-w64-mingw32/x64 (64-bit)(Windows >= 8 x64 (build 9200)).

 *  Function call: `makeDataReport(data = testData, file = "testdata_dateprob.rmd", 
    replace = TRUE)`