### 2021-10-07 latex/dataMaid_presidentData.Rmd

---
dataMaid: yes
title: Dirty president data
subtitle: "Autogenerated data summary from dataMaid"
date: 2018-04-11 01:02:26
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
```{r visualFunctions, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggAggHist <- getFromNamespace("ggAggHist", "dataMaid")
ggAggBarplot <- getFromNamespace("ggAggBarplot", "dataMaid")
```
# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations         46

Number of variables            11
---------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

----------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$             $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$

Identify case issues                                   $\times$    $\times$   $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$

Identify levels with < 6 obs.                                      $\times$   $\times$

Identify outliers                                                                        $\times$   $\times$             $\times$ 
----------------------------------------------------------------------------------------------------------------------------------

Non-supported variable types were set to be handled in the following way:

* Name is treated as character

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

--------------------------------------------------------------------------------------------------
&nbsp;                   Variable class     # unique values   Missing observations  Any problems? 
------------------------ ---------------- ----------------- ---------------------- ---------------
[lastName]               Name                            40                 0.00 %    $\times$

[firstName]              Name                            31                 0.00 %    $\times$

[orderOfPresidency]      factor                          46                 0.00 %    $\times$

[birthday]               Date                            45                 0.00 %    $\times$

[stateOfBirth]           character                       23                 0.00 %    $\times$

[assassinationAttempt]   numeric                          2                 0.00 %

[sex]                    factor                           1                 0.00 %    $\times$

[ethnicity]              factor                           2                 0.00 %    $\times$

[presidencyYears]        numeric                         11                 4.35 %    $\times$

[ageAtInauguration]      character                       23                 0.00 %    $\times$

[favoriteNumber]         complex                         11                 0.00 %    $\times$    
--------------------------------------------------------------------------------------------------

# Variable list
## lastName

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                  Name

Number of missing obs.      0 (0 %)

Number of unique values          40

Mode                        "Adams"
-----------------------------------

\emini
\bminitwo
```{r Var-1-lastName, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:40, .Label = c(" Truman", 
"Adams", "Arathornson", "Arthur", "Buchanan", "Bush", "Carter", 
"Cleveland", "Clinton", "Coolidge", "Eisenhower", "Fillmore", 
"Ford", "Garfield", "Grant", "Harding", "Harrison", "Hayes", 
"Hoover", "Jackson", "Jefferson", "Johnson", "Kennedy", "Lincoln", 
"Madison", "McKinley", "Monroe", "Nixon", "Obama", "Pierce", 
"Polk", "Reagan", "Roosevelt", "Taft", "Taylor", "Trump", "Tyler", 
"Van Buren", "Washington", "Wilson"), class = "factor"), y = c(1L, 
2L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L)), .Names = c("x", "y"), row.names = c(NA, 
-40L), class = "data.frame"), vnam = "lastName")
```
\emini

- The following values appear with prefixed or suffixed white space: \" Truman\".

\fullline

## firstName

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                  Name

Number of missing obs.      0 (0 %)

Number of unique values          31

Mode                        "James"
-----------------------------------

\emini
\bminitwo
```{r Var-2-firstName, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:31, .Label = c(".", 
"Abraham", "Andrew", "Aragorn", "Barack", "Benjamin", "Calvin", 
"Chester", "Dwight", "Franklin", "George", "Gerald", "Grover", 
"Harry", "Herbert", "James", "Jimmy", "John", "Lyndon", "Martin", 
"Millard", "Richard", "Ronald", "Rutherford", "Theodore", "Thomas", 
"Ulysses", "Warren", "William", "Woodrow", "Zachary"), class = "factor"), 
    y = c(1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 3L, 1L, 2L, 
    1L, 1L, 5L, 1L, 4L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    4L, 1L, 1L)), .Names = c("x", "y"), row.names = c(NA, -31L
), class = "data.frame"), vnam = "firstName")
```
\emini

- The following suspected missing value codes enter as regular values: \".\".

\fullline

## orderOfPresidency

* The variable is a key (distinct values for each observation).

\fullline

## birthday

\bminione

--------------------------------------------------
Feature                                     Result
------------------------- ------------------------
Variable type                                 Date

Number of missing obs.                     0 (0 %)

Number of unique values                         45

Mode                                  "1837-03-18"

Min. and max.               1300-03-01; 1961-08-04

1st and 3rd quartiles       1790-03-29; 1890-10-14
--------------------------------------------------

\emini
\bminitwo
```{r Var-4-birthday, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("1300-03-01", 
"1333-03-27", "1366-04-22", "1399-05-18", "1432-06-13", "1465-07-09", 
"1498-08-04", "1531-08-31", "1564-09-25", "1597-10-21", "1630-11-16", 
"1663-12-13", "1697-01-07", "1730-02-03", "1763-03-01", "1796-03-26", 
"1829-04-22", "1862-05-18", "1895-06-13", "1928-07-09"), class = "factor"), 
    Freq = c(1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
    0L, 5L, 8L, 6L, 9L, 6L, 7L, 4L), xmin = structure(c(-244653, 
    -232574, -220495, -208416, -196337, -184258, -172179, -160100, 
    -148021, -135942, -123863, -111783, -99704, -87625, -75546, 
    -63467, -51388, -39309, -27230, -15151), class = "Date"), 
    xmax = structure(c(-232574, -220495, -208416, -196337, -184258, 
    -172179, -160100, -148021, -135942, -123863, -111783, -99704, 
    -87625, -75546, -63467, -51388, -39309, -27230, -15151, -3072
    ), class = "Date"), ymin = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
    0, 0, 0, 0, 0, 0, 0, 0, 0, 0), ymax = c(1L, 0L, 0L, 0L, 0L, 
    0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 5L, 8L, 6L, 9L, 6L, 7L, 4L
    )), .Names = c("factorV", "Freq", "xmin", "xmax", "ymin", 
"ymax"), row.names = c(NA, -20L), class = "data.frame"), vnam = "birthday")
```
\emini

- Note that the following possible outlier values were detected: \"1300-03-01\".

\fullline

## stateOfBirth

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            23

Mode                           "Ohio"
-------------------------------------

\emini
\bminitwo
```{r Var-5-stateOfBirth, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:23, .Label = c("Arkansas", 
"California", "Connecticut", "Georgia", "Gondor", "Hawaii", "Illinois", 
"Iowa", "Kentucky", "Massachusetts", "Missouri", "Nebraska", 
"New Hampshire", "New Jersey", "New York", "New york", "North Carolina", 
"Ohio", "Pennsylvania", "Tennessee", "Texas", "Vermont", "Virginia"
), class = "factor"), y = c(1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
4L, 1L, 1L, 1L, 2L, 4L, 1L, 2L, 8L, 1L, 1L, 2L, 2L, 7L)), .Names = c("x", 
"y"), row.names = c(NA, -23L), class = "data.frame"), vnam = "stateOfBirth")
```
\emini

- Note that there might be case problems with the following levels: \"New York\", \"New york\".

\fullline

## assassinationAttempt

* Note that this variable is treated as a factor variable below, as it only takes a few unique values.

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               numeric

Number of missing obs.      0 (0 %)

Number of unique values           2

Mode                            "0"
-----------------------------------

\emini
\bminitwo
```{r Var-6-assassinationAttempt, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("0", 
"1"), class = "factor"), y = c(35L, 11L)), .Names = c("x", "y"
), row.names = c(NA, -2L), class = "data.frame"), vnam = "assassinationAttempt")
```
\emini

\fullline

## sex

* The variable only takes one (non-missing) value: \"Male\". The variable contains 0 \% missing observations.

\fullline

## ethnicity

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                    factor

Number of missing obs.          0 (0 %)

Number of unique values               2

Mode                        "Caucasian"
---------------------------------------

\emini
\bminitwo
```{r Var-8-ethnicity, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("African American", 
"Caucasian"), class = "factor"), y = c(1L, 45L)), .Names = c("x", 
"y"), row.names = c(NA, -2L), class = "data.frame"), vnam = "ethnicity")
```
\emini

- Note that the following levels have at most five observations: \"African American\".

\fullline

## presidencyYears

\bminione

--------------------------------------
Feature                         Result
------------------------- ------------
Variable type                  numeric

Number of missing obs.      2 (4.35 %)

Number of unique values             10

Median                               4

1st and 3rd quartiles          3.75; 8

Min. and max.                   0; Inf
--------------------------------------

\emini
\bminitwo
```{r Var-9-presidencyYears, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("[0,0.6]", 
"(0.6,1.2]", "(1.2,1.8]", "(1.8,2.4]", "(2.4,3]", "(3,3.6]", 
"(3.6,4.2]", "(4.2,4.8]", "(4.8,5.4]", "(5.4,6]", "(6,6.6]", 
"(6.6,7.2]", "(7.2,7.8]", "(7.8,8.4]", "(8.4,9]", "(9,9.6]", 
"(9.6,10.2]", "(10.2,10.8]", "(10.8,11.4]", "(11.4,12]"), class = "factor"), 
    Freq = c(2L, 1L, 0L, 4L, 4L, 0L, 15L, 0L, 3L, 0L, 0L, 3L, 
    0L, 10L, 0L, 0L, 0L, 0L, 0L, 1L), xmin = c(0, 0.6, 1.2, 1.8, 
    2.4, 3, 3.6, 4.2, 4.8, 5.4, 6, 6.6, 7.2, 7.8, 8.4, 9, 9.6, 
    10.2, 10.8, 11.4), xmax = c(0.6, 1.2, 1.8, 2.4, 3, 3.6, 4.2, 
    4.8, 5.4, 6, 6.6, 7.2, 7.8, 8.4, 9, 9.6, 10.2, 10.8, 11.4, 
    12), ymin = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
    0, 0, 0, 0, 0), ymax = c(2L, 1L, 0L, 4L, 4L, 0L, 15L, 0L, 
    3L, 0L, 0L, 3L, 0L, 10L, 0L, 0L, 0L, 0L, 0L, 1L)), .Names = c("factorV", 
"Freq", "xmin", "xmax", "ymin", "ymax"), row.names = c(NA, -20L
), class = "data.frame"), vnam = "presidencyYears")
```
\emini

- The following suspected missing value codes enter as regular values: \"Inf\".

- Note that the following possible outlier values were detected: \"0\", \"1\", \"Inf\".

\fullline

## ageAtInauguration

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            23

Mode                             "54"
-------------------------------------

\emini
\bminitwo
```{r Var-10-ageAtInauguration, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:23, .Label = c("42", 
"43", "46", "47", "48", "49", "50", "51", "52", "54", "55", "56", 
"57", "58", "60", "61", "62", "64", "65", "68", "69", "70", "87"
), class = "factor"), y = c(1L, 1L, 2L, 2L, 1L, 2L, 2L, 4L, 2L, 
5L, 4L, 3L, 4L, 1L, 1L, 3L, 1L, 2L, 1L, 1L, 1L, 1L, 1L)), .Names = c("x", 
"y"), row.names = c(NA, -23L), class = "data.frame"), vnam = "ageAtInauguration")
```
\emini

- Note: The variable consists exclusively of numbers and takes a lot of different values. Is it perhaps a misclassified numeric variable?

\fullline

## favoriteNumber

* The variable has class complex which is not supported by dataMaid.

\fullline

Report generation information:

 *  Created by Claus Thorn Ekstrøm (username: `cld189`).

 *  Report creation time: Wed Apr 11 2018 01:02:26

 *  Report Was run from directory: `/Users/cld189/ku/R/cleanR/latex`

 *  dataMaid v1.1.0 [Pkg: 2018-03-02 from local (ekstroem/dataMaid@NA)]

 *  R version 3.4.4 (2018-03-15).

 *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.3).

 *  Function call: `makeDataReport(data = presidentData, replace = TRUE, checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")), 
    reportTitle = "Dirty president data", treatXasY = list(Name = "character"))`