### 2021-10-07 latex/dataMaid_toyData.Rmd

---
dataMaid: yes
title: toyData
subtitle: "Autogenerated data summary from dataMaid"
date: 2018-04-11 01:02:39
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations         15

Number of variables             6
---------------------------------

* Only variables that were deemed potentially problematic are included in this summary.

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

----------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$             $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$

Identify levels with < 6 obs.                          $\times$    $\times$   $\times$

Identify case issues                                   $\times$    $\times$   $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$

Identify outliers                                                                        $\times$   $\times$             $\times$ 
----------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

-----------------------------------------------------------------------------------------
&nbsp;          Variable class     # unique values   Missing observations  Any problems? 
--------------- ---------------- ----------------- ---------------------- ---------------
[pill]          factor                           3                13.33 %    $\times$

[events]        numeric                          7                26.67 %    $\times$

[region]        factor                           7                 0.00 %    $\times$

[change]        numeric                         15                 0.00 %    $\times$

[id]            factor                          15                 0.00 %    $\times$

[spotifysong]   factor                           1                 0.00 %    $\times$    
-----------------------------------------------------------------------------------------

# Variable list
## pill

- Note that the following levels have at most five observations: \"blue\".

\fullline

## events

- Note that the following possible outlier values were detected: \"82\".

\fullline

## region

- The following suspected missing value codes enter as regular values: \" \", \".\".

- The following values appear with prefixed or suffixed white space: \" \".

- Note that the following levels have at most five observations: \" \", \".\", \"OTHER\", \"a\", \"b\", \"c\", \"other\".

- Note that there might be case problems with the following levels: \"OTHER\", \"other\".

\fullline

## change

- Note that the following possible outlier values were detected: \"1.12\", \"1.51\", \"1.6\".

\fullline

## id

* The variable is a key (distinct values for each observation).

\fullline

## spotifysong

* The variable only takes one (non-missing) value: \"Irrelevant\". The variable contains 0 \% missing observations.

\fullline

Report generation information:

 *  Created by Claus Thorn Ekstrøm (username: `cld189`).

 *  Report creation time: Wed Apr 11 2018 01:02:39

 *  Report Was run from directory: `/Users/cld189/ku/R/cleanR/latex`

 *  dataMaid v1.1.0 [Pkg: 2018-03-02 from local (ekstroem/dataMaid@NA)]

 *  R version 3.4.4 (2018-03-15).

 *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.3).

 *  Function call: `makeDataReport(data = toyData, onlyProblematic = TRUE, mode = "check", 
    replace = TRUE)`